Doubly-Robust Inference in R using \\texttt{drtmle}

### 2022-12-29 paper/reviews/response.md

Reviewer 1

I have enjoyed reading about the authors’ software, which I believe has the potential to lead to more robust inferences in observational studies. The following are some relatively minor comments.

Sections 2.2, 2.3, and 2.4– GCOMP, IPTW and AIPTW are commonly used estimators in applications. To improve the readability of the manuscript, I would perhaps shorten and combine sections 2.2, 2.3 and 2.4 into just one section describing existing estimators.

_Response_: Thanks for the suggestion. We have shortened and combined these sections.

Section 3– Unless I am missing something, I would mention that, in addition to the specific moment conditions described by the authors, the estimators need to satisfy also additional and untestable small-bias requirements in order to achieve doubly-robust asymptotic linearity. For example, the second-order terms in Appendix B in Benkeser et al (2017) need to be asymptotically negligible.

_Response_: Thanks, we have added references to the exact regularity conditions as they appear in the original manuscripts. We have also added the sentence

	> The regularity conditions required to achieve asymptotic linearity include assumptions on the convergence rates of the OR, PS, R-OR, and R-PS to their true counterparts."

Page 6, line starting with “The key theory…”– Should the estimators of the OR and PS satisfy equation (5) rather than (4)?

Page 7, line starting with “Further, they showed…”– Should the estimators of the OR, PS, R-OR satisfy equations (8)-(9) instead of (5)-(6)?

_Response_: Thanks for pointing out. There was some Latex/bookdown referencing gore happening with the original submission. We appreciate the reviewers catching this. It has been corrected in the revision.

Because of the crucial role that the reduced univariate regressions play in obtaining doubly-robust inference, I wonder whether it could be useful to have a simple way of checking the model fit for their estimators. For instance, would it make sense to have functions that can plot the residuals and fitted values against the estimated covariate?

_Response_: The package includes a `plot` method that can be used towards this end. We refer to this method in Section 4.1.1.

Reviewer 2

In this software tutorial, the authors detail and implement state-of-the-art nonparametric estimators of counterfactual means, or the average treatment effect (ATE), from observational studies, together with valid doubly robust statistical inferential tools (confidence intervals and hypothesis tests). Standard inference for doubly robust estimators of the ATE (e.g., AIPW or TML estimators) that employ flexible nuisance estimators is not strictly doubly robust in the same sense as the estimators themselves, i.e., consistency of the estimated ATE only requires consistency of one of two nuisance functions, but this robustness does not apply to asymptotic linearity/normality. That said, valid doubly robust inferential procedures have been proposed by van der Laan (2014) and Benkeser et al. (2017), in which additional nuisance models must be fit and satisfy certain estimating equations. However, these procedures are challenging and involved, thus the drtmle software package fills an important gap by implementing these methods. Beyond these approaches, which extend the usual TML estimator (drtmle function), the authors also include extensions to cross-validated TMLE (cvFolds option, i.e., using sample-splitting), and implement an IPW estimator with valid inference when using a flexible estimator for the propensity score (adaptive_iptw function).

The article does a good job of motivating and describing the underlying methodology well for readers familiar with doubly robust estimators of the ATE, but not with the doubly robust inference literature. Overall, the manuscript is also very clearly written, and the package seems to have a nice design. Specific comments are detailed below.

The workhorse functions (drtmle and adaptive_iptw) allow for missingness in treatment (A) and outcome (Y) values (e.g., with DeltaA and DeltaY control parameters within ‘glm_g’). This bivariate missingness setting doesn’t appear (unless I’m missing something) to have been formally considered in either van der Laan (2014) or Benkeser et al. (2017). I am concerned whether such missingness may lead to non-identifiability of the counterfactual mean—or “treatment-specific marginal mean” as referred to by the authors—functionals, without further assumptions. Are the authors implicitly invoking concrete missingness assumptions under which the implemented approaches target the original functional of interest (e.g., something like Delta_A x Delta_Y independent of (A, Y) given W)? Or are the methods simply standard, or used for convenience? It would enhance the paper if this issue were clarified.

_Response_: Thanks for the suggestion. We have added the formal assumptions needed for identification to that section (copied here):

if we define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively, then we can still perform valid inference on $\psi_0$ under the assumptions that (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$.

Related to point 1., why is missingness in the covariates (W) not considered? I am aware that this would be a very challenging scenario methodologically (and this is a good enough reason to not deal with it), but I am just curious whether there is something fundamentally different about W in this setting which precludes consideration of it being partially missing. Any guidance on what a practitioner should do if W is partially missing?

_Response_: Thanks for the suggestion. We have added a brief paragraph on missingness in $W$.

In section 4.2 on standard error estimation, the discussion is focused on the AIPW and standard TMLE (i.e., not the DR-TMLE implemented in the package). They say “The ideas generalize immediately to the DR-TML estimator, but the formulas are more complex in this latter case”. Presumably, that means that inference within drtmle is based on the empirical variance of the influence function of the DR-TMLE, but authors should elaborate at least a little. If the formulas are too complex to be presented within the manuscript, readers should be referred to the relevant sections of van der Laan (2014) and/or Benkeser et al. (2017).

_Response_: Thanks for the suggestion. We have restructured this section a bit and refer to the specific location of the formula in Benkeser et al (2017).

Several times, the equations (6) and (7) are referred to as if they were estimating equations, but it seems like they are definitions of new nuisance functions (at the “population” level).

a. On page 7, “Further, they showed that if estimators of the OR, PS, and R-OR satisfy equations (5)–(6)”. Equation (5) is empirical, but 6 is not. Should this say equations (8)—(9)?

b. On page 8, “Now, we have argued that a TML estimator may achieve doubly robust limiting behavior if it additionally satisfies equations (6) and (7)”. Should this say equations (9) and (10)?

c. On Page 8, “iteratively modify initial estimates of the OR and PS via the fluctuateQ() and fluctuateG() functions until equations (5)–(7) are approximately solved”. Should this say equations (8)—(10)?

_Response_: All equation referencing has been corrected. Sorry for the confusion.

Other specific minor comments:

Page 3, “relying upon Neymon orthogonality” should say Neyman?

_Response_: Correct. Thanks for catching.

Page 4, “Under additional assumptions, these parameters have interpretations…”. Should at least give a reference here to the causal assumptions.

_Response_: We have added a few short sentences on causal assumptions.

Page 7, equation (9), I think the Q^bar term in the numerator and g_n term in the denominator should each have an asterisk.

_Response_: Yes, thanks. Good catch.

Page 9, “r round(EY0, 2)” looks like unrendered markdown. What is the numerical value?

_Response_: This has been corrected.

Page 18, “methods that compute doubly robust confidence intervals”. Can this be clarified to “Wald-style” confidence intervals? It seems from context that these are just based directly on the standard error estimates from the previous section, but would be good to clarify in opening paragraph.

_Response_: Done

### 2022-12-29 paper/using_drtmle_obstudies.Rmd
updates to address reviewer comments

In tandem with growing interest in and usage of doubly robust estimation
approaches, open source software tools implementing these methods have
proliferated. Notably extant software packages include the \texttt{tmle}
\texttt{R} package [@gruber2012tmle], which implements TML estimators of the
effects of static single time point interventions; the \texttt{ltmle} \texttt{R}
package [@lendle2017ltmle], which implements TML estimators of the effects of
multiple time point interventions, subject to time-dependent confounding and
censoring, as well as estimators of the parameters of a longitudinal working
marginal structural model; the \texttt{tmle3} \texttt{R} package
[@coyle2021tmle3-rpkg], which serves as a generalizable engine for implementing
TML estimators within the \texttt{tlverse} ecosystem [@vdl2022targeted],
designed to provide a common and extensible framework for TML estimation of the
effects of static, dynamic, and stochastic single time point interventions; and
the \texttt{doubleml} \texttt{R} and \texttt{Python} packages
[@bach2021doublemlR; @bach2022doublemlPy], which implement estimators of causal
effects within the double machine learning framework [@chernozhukov2017double;
@chernozhukov2018double], reyling upon Neymon orthogonality and sample-splitting
for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
as, to our knowledge, the only implementation of statistical procedures designed
to attain doubly robust inference [@benkeser:2017:biometrika], uniquely
positioning \texttt{drtmle} as a robust tool for the estimation of treatment
effects in observational data settings.

@@

Suppose the observed data consist of $n$ independent and identically distributed
copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
and $P_0$ is the true data-generating distribution. The parameters of interest
are
\begin{equation}
  \psi_0(a) = E_0\{\overline{Q}(a,W)\} = \int \overline{Q}_0(a,w) dQ_{W,0}(w)
  \ , \ \mbox{for} \ a = 0,1 \ ,
  \label{parameterDef}
\end{equation}
where $\overline{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
assumptions, these parameters have interpretations as mean counterfactual
outcomes under the different treatments, with contrasts between these means
defining causal effects.

## G-computation estimators

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n
  \overline{Q}_n(a, W_i) \ .
  \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

## Inverse probability of treatment estimators

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\overline{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may also be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

## Doubly robust estimators

@@

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One prominent example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\}
  \overline{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators: It exhibits _double robustness_ in that it is consistent for
$\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
consistent for the true PS. Furthermore, if both the OR and PS are consistently
estimated and regularity conditions hold, one can establish a normal limiting
distribution for the AIPTW estimator. Closed-form estimators of the variance of
this limiting distribution can be derived that may be used to construct
Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

@@

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qoverliner} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
  \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @vdl:2014:ijb showed that if estimates of the
OR $\overline{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR
$\overline{Q}_{r,n}^{\star}$, and R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}(a, W_i)}
{g_n(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \overline{Q}_n^{\star}(a, W_i) \}
= 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
robust not only with respect to consistency, but also with respect to its
limiting distribution. The author additionally proposed a TMLE algorithm to map
initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
these key equations.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w) \} \ ,
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \overline{Q}_n(W) = \overline{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations
(\ref{meanEIFZero})--(\ref{eqn:qoverliner}), and if, additionally, estimators of
the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\overline{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\overline{Q}_n^{\star}$ also enjoys a
doubly robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly robust variance estimates.

>>>>

In tandem with growing interest in and usage of doubly robust estimation
approaches, open source software tools implementing these methods have
proliferated. Notably extant software packages include the \texttt{tmle}
\texttt{R} package [@gruber2012tmle], which implements TML estimators of the
effects of static single time point interventions; the \texttt{ltmle} \texttt{R}
package [@lendle2017ltmle], which implements TML estimators of the effects of
multiple time point interventions, subject to time-dependent confounding and
censoring, as well as estimators of the parameters of a longitudinal working
marginal structural model; the \texttt{tmle3} \texttt{R} package
[@coyle2021tmle3-rpkg], which serves as a generalizable engine for implementing
TML estimators within the \texttt{tlverse} ecosystem [@vdl2022targeted],
designed to provide a common and extensible framework for TML estimation of the
effects of static, dynamic, and stochastic single time point interventions; and
the \texttt{doubleml} \texttt{R} and \texttt{Python} packages
[@bach2021doublemlR; @bach2022doublemlPy], which implement estimators of causal
effects within the double machine learning framework [@chernozhukov2017double;
@chernozhukov2018double], reyling upon Neyman orthogonality and sample-splitting
for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
as, to our knowledge, the only implementation of statistical procedures designed
to attain doubly robust inference [@benkeser:2017:biometrika], uniquely
positioning \texttt{drtmle} as a robust tool for the estimation of treatment
effects in observational data settings.

@@

Suppose the observed data consist of $n$ independent and identically distributed
copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
and $P_0$ is the true data-generating distribution. The parameters of interest
are
\begin{equation}
  \psi_0(a) = E_0\{\overline{Q}(a,W)\} = \int \overline{Q}_0(a,w) dQ_{W,0}(w)
  \ , \ \mbox{for} \ a = 0,1 \ ,
  (\#eq:parameterDef)
\end{equation}
where $\overline{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level) treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$.

Under additional assumptions, these parameters identify the mean counterfactual
outcomes under the different treatments. Specifically, we denote by $Y(a)$ a well-defined 
counterfactual outcome that would be observed under treatment $a$.
If the conditional randomization assumption ($Y(a) \perp A \mid W$) and positivity
conditions ($Pr_0\{ Pr_0(A = a \mid W) > 0 \} = 1$) hold, then $\psi(a)$ identifies
the mean of $Y(a)$.

## Estimators

The G-computation estimator of $\psi_0(a)$ may be motivated directly by \@ref(eq:parameterDef).
Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into \@ref(eq:parameterDef),
$\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n
  \overline{Q}_n(a, W_i) \ .
  (\#eq:gcomp)
\end{equation}
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that \@ref(eq:gcomp) can be equivalently written as
\begin{align}
  E_0\{\overline{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  (\#eq:iptwID)
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
\@ref(eq:iptwID) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap. 
If adaptive approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing (as in @ertefaie2022nonparametric and @hejazi2022efficient).

@@

The augmented IPTW (AIPTW) estimator
\begin{equation*}
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\}
  \overline{Q}_n(a, W_i) \ ,
\end{equation*}
adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} 
  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
  \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n(a,W_i)\} \ . (\#eq:oneStepEst)
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. First, it exhibits _double robustness_ in that it is consistent for
$\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
consistent for the true PS. Furthermore, if both the OR and PS are consistently
estimated and regularity conditions hold, one can establish a normal limiting
distribution for the AIPTW estimator. Closed-form estimators of the variance of
this limiting distribution can be derived that may be used to construct
Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
\begin{equation}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ , (\#eq:meanEIFZero)
\end{equation}
then the GCOMP estimator based on $\overline{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies \@ref(eq:meanEIFZero). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

@@

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation \@ref(eq:meanEIFZero) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} (\#eq:qoverliner) \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
  (\#eq:grbivariate)
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @vdl:2014:ijb showed that if estimates of the
OR $\overline{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR
$\overline{Q}_{r,n}^{\star}$, and R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , (\#eq:tmleEIFSolve) \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}^{\star}(a, W_i)}
{g_n^{\star}(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
(\#eq:tmleAsolve) \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \overline{Q}_n^{\star}(a, W_i) \}
= 0 \ ,
(\#eq:tmleQsolve)
\end{align}
then, under regularity conditions described in Theorem 3 of @vdl:2014:ijb, 
the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
robust not only with respect to consistency, but also with respect to its
limiting distribution. The author additionally proposed a TMLE algorithm to map
initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
these key equations. The regularity conditions required to achieve asymptotic
linearity include assumptions on the convergence rates of the OR, PS, R-OR, and R-PS
to their true counterparts.

@benkeser:2017:biometrika demonstrated alternative equations that can be
satisfied to achieve this form of double-robustness. In particular, these
authors formulated an alternative version of the R-PS,
\begin{align*}
  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \overline{Q}_n(W) =
  \overline{Q}_n(w) \} \ , 
\end{align*}
which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR,
\begin{align*}
  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
  {g_n(a \mid W)} \ \bigg\vert \ \overline{Q}_n(W) = \overline{Q}_n(w) \biggr\} \ ,
\end{align*}
which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression.
@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
easier to estimate consistently than R-PS. Further, they showed that if
estimators of the OR, PS, and R-OR satisfy equations
\@ref(eq:tmleEIFSolve)--\@ref(eq:tmleAsolve), and if, additionally, estimators of
the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
\begin{equation*}
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
\overline{Q}_n^{\star}(a,W_i)\} = 0 \ ,
\end{equation*}
then the GCOMP estimator based on this $\overline{Q}_n^{\star}$ also enjoys a
doubly robust limiting distribution. @benkeser:2017:biometrika provided a
TMLE algorithm that produced such estimates and provided closed-form,
doubly robust variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and, practically, the two estimators are generally seen as competitors
for estimating marginal means. In equation \@ref(eq:oneStepEst), we were able to
generate a doubly robust estimator by adding a term to the GCOMP estimator. This
term is exactly the left-hand side of the key equation \@ref(eq:meanEIFZero), the
key equation
that is solved by the standard TMLE. This is no coincidence -- this term is
crucial in studying the asymptotic properties of the GCOMP, AIPTW, and TML
estimators (for discussion, see @benkeser:2017:biometrika, Section 2.2). Recall
that a TML estimator achieves doubly-robust consistency by satisfying equation
\@ref(eq:meanEIFZero), while the AIPTW achieves doubly robust consistency by
combining the GCOMP estimator with the left-hand side of this term. Now, we have
argued that a TML estimator may achieve doubly robust limiting behavior if it
additionally satisfies equations \@ref(eq:tmleAsolve) and
\@ref(eq:tmleQsolve). It is thus sensible to wonder whether an AIPTW-style
estimator could be constructed by combining the left-hand side of these two
additional equations with the AIPTW estimator. @benkeser:2017:biometrika showed
that, surprisingly, this AIPTW-style estimator does not achieve the same
doubly robust properties as the TML estimator. Thus, while AIPTW and TMLE are
generally competitors for producing estimators doubly robust with respect to
consistency, for doubly robust inference, TML estimators are preferable.
